Qubit tests from classical codes

### 2022-12-31 quantum_codes_test.tex
some comments, many questions

  A \emph{tracial von Neumann algebra} is a pair $(\mM,\tau)$ of a von Neumann algebra $\mM$ together with a normal faithful tracial state $\tau$ on $\mM$, which we often refer to as the \emph{trace}. The main example of interest is $\mM=M_n(\C)$, the algebra $n\times n$ complex matrices, with $\tau$ the dimension-normalized trace, which we denote $\tr(M)=\frac{1}{d}\Tr(M)$. 	We write $\|x\|_2=\tau(x^*x)^{1/2}$ for the $2$-norm on $\mM$.

	We denote $\mM_\infty = \mM \overline{\otimes} B(\ell_2)$ with (infinite) trace $\tau_\infty = \tau \otimes \Tr$. We identify $\mM$ with $\mM\otimes e_{1,1}\subset \mM_\infty$.

	We let $\field$ denote the field with two elements. For $u\in \F^n$ for some $n$, we write $|u|$ for the Hamming weight of $u$, i.e.\ the number of nonzero coordinates.

@@

\begin{proof}
If $\eps\geq \frac{1}{2}$ the conclusion is trivial (for a suitably large implicit constant), so assume $\eps<\frac{1}{2}$. 
Define 
\[\tilde{Q}_i = w^* P_i w  + \frac{1}{|\mI|}\big(1_\mM - w^* w\big) \in \mM\;.\]
Then $\{\tilde{Q}_i\}$ is a POVM on $\mM$. Moreover, 
\begin{align*}
\sum_{i} \tau\big( (\tilde{Q}_i)^2 \big) &\geq \sum_{i} \tau\big( \big(w^* P_i w \big)^2 \big) \\
&= \sum_{i} \tau\big(  w^* P_i w w^*P_i w \big)\\
&= \sum_i \tau\big(  w^* P_i  P P_i w \big) - \sum_i \tau\big( w^* P_i  ( P - w w^*) P_i w \big)\\
&\geq 1 - \eps -  \tau_\infty\Big(\big( P - w w^*\big)\Big(\sum_i  P_i w w^* P_i\Big)\Big)\\ 
&\geq 1- \eps- \tau_\infty\big( P- w w^*\big)\;,
\end{align*}
where the fourth line uses that $P_iPP_i=P_i$, $\sum_i P_i = \Id_\mN$ and the definition of $\eps$ for the first term, and cyclicity of the trace for the second, and the last uses $\|ww^*\|,\|\sum_i P_i\|_\infty\leq 1$. By assumption, 
\begin{align*}
\tau_\infty\big( P- w w^*\big) \,\leq\, \eps\, \tau_\infty(P)\,\leq \frac{\eps}{1-\eps}\;.
\end{align*}
where the last inequality is because by definition, $\tau^N(P)=1$, thus
\[1-\eps \,\leq\, \tau^\mN(ww^*) \,=\, \frac{\tau_\infty(ww^*)}{\tau_\infty(P)}\,=\, \frac{\tau^\mM(ww^*)}{\tau_\infty(P)}  \,\leq\, \frac{1}{\tau_\infty(P)}\]
since $\tau^\mM$ is normalized and $ww^*\leq 1_\mM$. Overall, 
\[ \sum_{i} \tau\big( (\tilde{Q}_i)^2 \big) \,\geq\, 1-\eps-\frac{\eps}{1-\eps}\,\geq\, 1-3\eps\;.\]
To conclude we apply~\cite[Theorem 1.2]{de2021orthogonalization} to obtain a projective measurement $\{Q_i\}$ on $\mM$ such that 
\begin{equation*}
\sum_i \big\|{Q}_i - \tilde{Q}_i \big\|^2_2 \,=\, O(\eps)\;.
\end{equation*}
Finally,
\begin{align*}
\sum_i \big\|{Q}_i - w^*{P}_i w\big\|^2_2 &= \sum_i \Big\|{Q}_i - \tilde{Q}_i  + \frac{1}{|\mI|}\big(1_\mM - w^* w\big) \Big\|^2_2\\
&\leq  \sum_i 2\big\|{Q}_i - \tilde{Q}_i\big\|_2^2  + 2\frac{1}{|I|}\big\|1_\mM - w^* w\big\|_2^2 \\
&=  O(\eps)\;,
\end{align*}
where the second line is by the triangle inequality. 
\end{proof}

@@

\begin{definition}[Code]
A $[n,k,d]$ linear code is a $k$-dimensional subspace $\code\subseteq \field^n$ such that for every $u\in \code$ such that $u\neq 0$, $|u|\geq d$. Equivalently, we think of $\code$ as a linear map $\code:\field^k\to \field^n$. We write $G_\code \in \field^{k\times n}$ for the generator matrix defined by $ G_\code^T e_i =\code(e_i)$ for all $i\in\{1,\ldots,k\}$, where $\{e_1,\ldots,e_k\}$ is the standard basis of $\field^k$. 
\end{definition}

@@

\begin{definition}[Codeword test]
Let $\code$ be an $[n,k,d]$ linear code, $\delta:[0,1]\to[0,1]$, and $r\in \N$.
A $r$-local $\delta$-tester for $\code$ is a non-adaptive randomized oracle machine $M$ that makes at most $r$ queries to its proof tape and satisfies the following conditions:
\begin{itemize} 
\item (Completeness:) For any $u\in \code$, $\Pr_\rand( M(\rand,u)=1)=1$.
\item (Soundness:) For any $\eps\geq 0$ and any $u\in \field^n$ that is within (Hamming) distance at least  $\eps$ from $\code$, $\Pr_\rand(M(\rand,u)=0)\geq \delta(\eps)$. 
\end{itemize}
\end{definition}

@@

\begin{definition}[$\eps$-local presentation]
Let $\code$ be an $[n,k,d]$ linear code and $M$ an $r$-local tester for $\code$. An \emph{$\eps$-local presentation of $(\code,M)$} is a collection of projective measurements $\{A^i_a\}_{a\in\field} \subseteq\mM$ for $i\in\{1,\ldots,n\}$ such that for any $S\subseteq\{1,\ldots,n\}$ of size $|S|\leq r$ there are $S$ pairwise commuting projective measurements $\{B^{S,i}_a\}_{i\in S}$ and $\eps_S\geq 0$ such that 
\begin{itemize}
\item $\sum_{a\in \F^S} 1_{M(S,a)=1} \prod_{i\in S} B^{S,i}_{a_i}= \Id$
\item $\{A^i\}_{i\in S}$ and $\{B^{S,i}\}_{i\in S}$ are $\eps_S$-close
\end{itemize} 
and moreover $\Es{S\leftarrow M(\cdot)} \eps_S \leq \eps$. 
\end{definition}

@@

\begin{definition}[Synchronous strategy]
If $G=(\mX,\mu,\mA,D)$ is a game and $(\mM,\tau)$ a tracial von Neumann algebra, a \emph{synchronous strategy $\strategy$ for $G$ on $(\mM,\tau)$} is, for every $x\in \mX$, a projective measurement $(P^x_a)_{a\in \mA(x)}$ on $\mM$. The value of a strategy $\strategy$ in $G$ is 
\[ \omega(G;\strategy)\,=\, \sum_{(x,y)\in\mX\times\mX}\frac{1}{2}\big(\mu(x,y)+\mu(y,x)\big) \sum_{(a,b)\in\mA(x)\times\mA(y)} D(x,y,a,b)\, \tau\big(P^x_a \,P^y_b\big) \;.\footnote{Note the symmetrization of $\mu$. This is to avoid explicitly requiring $\mu$ to be permutation-invariant in the definition of a game.}\]
We say that $\strategy$ is \emph{perfect} if $\omega(G;\strategy)=1$.
\end{definition}

@@

\begin{proposition}\label{prop:sound-game}
Let $\code$ be an $[n,k,d]$ linear code and $M$ an $r$-local tester for $\code$. Let $\delta,\delta':[0,1]\to[0,1]$. Then the following hold:
\begin{enumerate}
\item If $M$ has quantum soundness $\delta$ then $G_{\code,M}$ is $(\delta'',\nu)$-robust, for some $\delta''(\eps)=O({\delta(2\eps)})$ and $\nu$ the uniform distribution on $\{1,\ldots,n\}\subseteq \mX$. 
\item If $G_{\code,M}$ is $(\delta',\nu)$-robust for $\nu$ the uniform distribution on $\{1,\ldots,n\}\subseteq \mX$, and $(\mC,M)$ is Abelian, then $M$ has quantum soundness $\delta''$, for some $\delta''(\eps) = O(\delta'(\sqrt{\eps}))$. 
\end{enumerate}
\end{proposition}

@@

\begin{definition}
Let $k\in \N$ and $\delta:[0,1]\to\R_+$. 
A $(k,\delta(\eps))$-qubit test is a synchronous game $G=(\mX,\mu,\mA,D)$ such there are two sets $S_X,S_Z\subseteq \field^k$ and an injection $\phi:(\{X\}\times S_Z) \cup (\{Z\}\times S_Z) \to \mX$ such that $\mA(\phi({X},a))=\mA(\phi({Z},b))=\field$ for all $a\in S_X$, $b\in S_Z$ and such that the following holds:
\begin{itemize}
\item (Completeness:) There is a synchronous strategy for $G$ on $\mM=M_{2^{k}}(\C)$ that succeeds with probability $1$ in $G$ and is such that $\widehat{P^{\phi({W},a)}} = \tau^W(a)$ for every $W\in\{X,Z\}$ and $a\in S_W$.
\item (Soundness:) Any synchronous strategy in $(\mM,\tau)$ for $G$ that succeeds with probability $1-\eps$ for some $\eps\geq 0$ is $\delta(\eps)$-close to a strategy on an algebra of the form $(M_{2^{k}}(\C)\otimes \mN,\tr\otimes \tau')$ and such that
\[\widehat{P^{\phi({W},a)}} = \tau^W(a)\otimes \Id_\mN\;.\]
\end{itemize}
\end{definition}

@@

\underline{Soundness:} Next we show soundness. Let $\strategy$ be a synchronous strategy for $G$  in  $(\mM,\tau)$ that succeeds with probability at least $1-\eps$. For $W\in\{X,Z\}$ let $\strategy^W$ be the strategy in $G_{\code,M}$ that is obtained by restricting $\strategy$ to the relevant measurements, i.e.\ the $P^{W,S}$ and $P^{W,i}$.  Then $\strategy^W$ succeeds with probability at least $1-3\eps$ in $G_{\code,M}$. Using the assumption that $M$ is $\delta(\cdot)$-quantum sound and the first item from Proposition~\ref{prop:sound-game} it follows that there is a $\delta_1 = O({\delta(6\eps)})$ such that for each $W\in\{X,Z\}$, $\strategy^W$ is $(\delta_1,\nu)$ close to a perfect strategy $\tilde{\strategy}^W$ on $(\mN^W,\tau^W)$, where $\nu$ is uniform on $\{1,\ldots,n\}$. This strategy has measurement operators $\{ \tilde{P}^{W,S}_a\}$ and $\{\tilde{P}^{W,i}_b\}$ associated with questions of the form $S$ and $i$ in $G_{\code,M}$. Using the definition of $\nu$, these satisfy that 
\begin{equation}\label{eq:main-0}
\Es{i\in\{1,\ldots,n\}}\sum_b \big\|P^{W,i}_b - (w^W)^* \tilde{P}_i^W w^W \big\|^2_2 \,=\, O(\delta_1)\;.
\end{equation}
Furthermore, since $(\code,M)$ is Abelian, there is a  POVM $\{\tilde{P}^W_u\}_{u\in \field^n}$ such that $\sum_{u\in \code}\tilde{P}^W_u = \Id$ and for each $i\in \{1,\ldots,n\}$, $\tilde{P}^{W,i}_b = \sum_{a:a_i=b} \tilde{P}^W_a$.% where  $u_i = G_\code e_i$.

>>>>

  A \emph{tracial von Neumann algebra} is a pair $(\mM,\tau)$ of a von Neumann algebra $\mM$ together with a normal faithful tracial state $\tau$ on $\mM$, which we often refer to as the \emph{trace}. The main example of interest is $\mM=M_n(\C)$, the algebra $n\times n$ complex matrices, with $\tau$ the dimension-normalized trace, which we denote $\tr(M)=\frac{1}{n}\Tr(M)$ \hnote{was $\frac{1}{d}$}. 	We write $\|x\|_2=\tau(x^*x)^{1/2}$ for the $2$-norm on $\mM$.

	We denote $\mM_\infty = \mM \overline{\otimes} B(\ell_2)$ \hnote{what do $\overline{\otimes}$ and $B(\ell_2)$ mean?} with (infinite) trace $\tau_\infty = \tau \otimes \Tr$ \hnote{what does infinite trace mean? also, should it be $\Tr$ or $\tr$?}. We identify $\mM$ with $\mM\otimes e_{1,1}\subset \mM_\infty$.

	We let $\field$ denote the field with two elements. For $u\in \F^n$ for some $n$, we write $|u|$ for the Hamming weight of $u$, i.e.\ the number of nonzero coordinates. \hnote{added:} For $a,b \in \field^k$, we write $a \cdot b$ to denote the inner product $\sum_{i=1}^k a_i b_i$.

@@

\begin{proof}
If $\eps\geq \frac{1}{2}$ the conclusion is trivial (for a suitably large implicit constant), so assume $\eps<\frac{1}{2}$. 
Define 
\[\tilde{Q}_i = w^* P_i w  + \frac{1}{|\mI|}\big(1_\mM - w^* w\big) \in \mM\;.\]
Then $\{\tilde{Q}_i\}$ is a POVM on $\mM$. Moreover, 
\begin{align*}
\sum_{i} \tau\big( (\tilde{Q}_i)^2 \big) &\geq \sum_{i} \tau\big( \big(w^* P_i w \big)^2 \big) \\
&= \sum_{i} \tau\big(  w^* P_i w w^*P_i w \big)\\
&= \sum_i \tau\big(  w^* P_i  P P_i w \big) - \sum_i \tau\big( w^* P_i  ( P - w w^*) P_i w \big)\\
&\geq 1 - \eps -  \tau_\infty\Big(\big( P - w w^*\big)\Big(\sum_i  P_i w w^* P_i\Big)\Big)\\ 
&\geq 1- \eps- \tau_\infty\big( P- w w^*\big)\;,
\end{align*}
where the fourth line uses that $P_iPP_i=P_i$, $\sum_i P_i = \Id_\mN$ and the definition of $\eps$ for the first term, and cyclicity of the trace for the second, and the last uses $\|ww^*\|,\|\sum_i P_i\|_\infty\leq 1$. By assumption, 
\begin{align*}
\tau_\infty\big( P- w w^*\big) \,\leq\, \eps\, \tau_\infty(P)\,\leq \frac{\eps}{1-\eps}\;.
\end{align*}
where the last inequality is because by definition, $\tau^N(P)=1$, thus
\[1-\eps \,\leq\, \tau^\mN(ww^*) \,=\, \frac{\tau_\infty(ww^*)}{\tau_\infty(P)}\,=\, \frac{\tau^\mM(ww^*)}{\tau_\infty(P)}  \,\leq\, \frac{1}{\tau_\infty(P)}\]
\hnote{why is $ww^* \leq 1_\mM$? The input to $w^*$ is from $P$, which belongs to ${\mM}_{\infty}$}
since $\tau^\mM$ is normalized and $ww^*\leq 1_\mM$. Overall, 
\[ \sum_{i} \tau\big( (\tilde{Q}_i)^2 \big) \,\geq\, 1-\eps-\frac{\eps}{1-\eps}\,\geq\, 1-3\eps\;.\]
To conclude we apply~\cite[Theorem 1.2]{de2021orthogonalization} to obtain a projective measurement $\{Q_i\}$ on $\mM$ such that 
\begin{equation*}
\sum_i \big\|{Q}_i - \tilde{Q}_i \big\|^2_2 \,=\, O(\eps)\;.
\end{equation*}
Finally,
\begin{align*}
\sum_i \big\|{Q}_i - w^*{P}_i w\big\|^2_2 &= \sum_i \Big\|{Q}_i - \tilde{Q}_i  + \frac{1}{|\mI|}\big(1_\mM - w^* w\big) \Big\|^2_2\\
&\leq  \sum_i 2\big\|{Q}_i - \tilde{Q}_i\big\|_2^2  + 2\frac{1}{|I|}\big\|1_\mM - w^* w\big\|_2^2 \\
&=  O(\eps)\;,
\end{align*}
where the second line is by the triangle inequality. 
\end{proof}

@@

\begin{definition}[Code]
A $[n,k,d]$ linear code \hnote{need to specify $\field$?} is a $k$-dimensional subspace $\code\subseteq \field^n$ such that for every $u\in \code$ such that $u\neq 0$, $|u|\geq d$. Equivalently, we think of $\code$ as a linear map $\code:\field^k\to \field^n$. We write $G_\code \in \field^{k\times n}$ for the generator matrix defined by $ G_\code^T e_i =\code(e_i)$ for all $i\in\{1,\ldots,k\}$, where $\{e_1,\ldots,e_k\}$ is the standard basis of $\field^k$. 
\end{definition}

@@

\begin{definition}[Codeword test]
Let $\code$ be an $[n,k,d]$ linear code, $\delta:[0,1]\to[0,1]$, and $r\in \N$.
A $r$-local $\delta$-tester for $\code$ is a \hnote{if we're using this Turing machine formalism, why restrict to non-adaptive?} non-adaptive randomized oracle machine $M$ that makes at most $r$ queries to its proof tape and satisfies the following conditions:
\begin{itemize} 
\item (Completeness:) For any $u\in \code$, $\Pr_\rand( M(\rand,u)=1)=1$.
\item (Soundness:) For any $\eps\geq 0$ and any $u\in \field^n$ that is within (Hamming) distance at least  $\eps$ from $\code$, $\Pr_\rand(M(\rand,u)=0)\geq \delta(\eps)$. 
\end{itemize}
\end{definition}

@@

\begin{definition}[$\eps$-local presentation]
Let $\code$ be an $[n,k,d]$ linear code and $M$ an $r$-local tester for $\code$. An \emph{$\eps$-local presentation of $(\code,M)$} is a collection of projective measurements $\{A^i_a\}_{a\in\field} \subseteq\mM$ for $i\in\{1,\ldots,n\}$ such that for any $S\subseteq\{1,\ldots,n\}$ of size $|S|\leq r$ there are $|S|$ pairwise commuting projective measurements $\{B^{S,i}_a\}_{i\in S}$ \hnote{the output alphabet is confusing here; here it seems like $a \in \field$ whereas below $a \in \F^S$.} and $\eps_S\geq 0$ such that 
\begin{itemize}
\item $\sum_{a\in \F^S} 1_{M(S,a)=1} \prod_{i\in S} B^{S,i}_{a_i}= \Id$
\item $\{A^i\}_{i\in S}$ and $\{B^{S,i}\}_{i\in S}$ are $\eps_S$-close
\end{itemize} 
and moreover $\Es{S\leftarrow M(\cdot)} \eps_S \leq \eps$. \hnote{The $1_{M(S,a)=1}$ notation conflicts with the $1_{\mM}$ notation above. Also sometimes $I$ is used, othertimes $1$ is used.}
\end{definition}

@@

\begin{definition}[Synchronous strategy]
If $G=(\mX,\mu,\mA,D)$ is a game and $(\mM,\tau)$ a tracial von Neumann algebra, a \emph{synchronous strategy $\strategy$ for $G$ on $(\mM,\tau)$} is, for every $x\in \mX$, a projective measurement $(P^x_a)_{a\in \mA(x)}$ on $\mM$. The value of a strategy $\strategy$ in $G$ is 
\[ \omega(G;\strategy)\,=\, \sum_{(x,y)\in\mX\times\mX}\frac{1}{2}\big(\mu(x,y)+\mu(y,x)\big) \sum_{(a,b)\in\mA(x)\times\mA(y)} D(x,y,a,b)\, \tau\big(P^x_a \,P^y_b\big) \;.\footnote{Note the symmetrization of $\mu$. This is to avoid explicitly requiring $\mu$ to be permutation-invariant in the definition of a game. \hnote{Didn't quite understand... what would go wrong without this symmetrization?}}\]
We say that $\strategy$ is \emph{perfect} if $\omega(G;\strategy)=1$.
\end{definition}

@@

\begin{proposition}\label{prop:sound-game}
Let $\code$ be an $[n,k,d]$ linear code and $M$ an $r$-local tester for $\code$. Let $\delta,\delta':[0,1]\to[0,1]$. Then the following hold:
\begin{enumerate}
\item If $M$ has quantum soundness $\delta$ then $G_{\code,M}$ \hnote{was this object defined yet?} is $(\delta'',\nu)$-robust, for some $\delta''(\eps)=O({\delta(2\eps)})$ and $\nu$ the uniform distribution on $\{1,\ldots,n\}\subseteq \mX$. 
\item If $G_{\code,M}$ is $(\delta',\nu)$-robust for $\nu$ the uniform distribution on $\{1,\ldots,n\}\subseteq \mX$, and $(\mC,M)$ is Abelian, then $M$ has quantum soundness $\delta''$, for some $\delta''(\eps) = O(\delta'(\sqrt{\eps}))$. 
\end{enumerate}
\end{proposition}

@@

\begin{definition}
Let $k\in \N$ and $\delta:[0,1]\to\R_+$. 
A $(k,\delta(\eps))$-qubit test is a synchronous game $G=(\mX,\mu,\mA,D)$ such there are two sets $S_X,S_Z\subseteq \field^k$ and an injection $\phi:(\{X\}\times S_Z) \cup (\{Z\}\times S_Z) \to \mX$ such that $\mA(\phi({X},a))=\mA(\phi({Z},b))=\field$ for all $a\in S_X$, $b\in S_Z$ and such that the following holds:
\begin{itemize}
\item (Completeness:) There is a synchronous strategy for $G$ on $\mM=M_{2^{k}}(\C)$ that succeeds with probability $1$ in $G$ and is such that $\widehat{P^{\phi({W},a)}} = \tau^W(a)$ \hnote{overloading $\tau$} for every $W\in\{X,Z\}$ and $a\in S_W$.
\item (Soundness:) Any synchronous strategy in $(\mM,\tau)$ for $G$ that succeeds with probability $1-\eps$ for some $\eps\geq 0$ is $\delta(\eps)$-close to a strategy on an algebra of the form $(M_{2^{k}}(\C)\otimes \mN,\tr\otimes \tau')$ and such that
\[\widehat{P^{\phi({W},a)}} = \tau^W(a)\otimes \Id_\mN\;.\]
\end{itemize}
\end{definition}

@@

\underline{Soundness:} Next we show soundness. Let $\strategy$ be a synchronous strategy for $G$  in  $(\mM,\tau)$ that succeeds with probability at least $1-\eps$. For $W\in\{X,Z\}$ let $\strategy^W$ be the strategy in $G_{\code,M}$ that is obtained by restricting $\strategy$ to the relevant measurements, i.e.\ the $P^{W,S}$ and $P^{W,i}$.  Then $\strategy^W$ succeeds with probability at least $1-3\eps$ in $G_{\code,M}$. Using the assumption that $M$ is $\delta(\cdot)$-quantum sound and the first item from Proposition~\ref{prop:sound-game} it follows that there is a $\delta_1 = O({\delta(6\eps)})$ such that for each $W\in\{X,Z\}$, $\strategy^W$ is $(\delta_1,\nu)$ close to a perfect strategy $\tilde{\strategy}^W$ on $(\mN^W,\tau^W)$, where $\nu$ is uniform on $\{1,\ldots,n\}$. This strategy has measurement operators $\{ \tilde{P}^{W,S}_a\}$ and $\{\tilde{P}^{W,i}_b\}$ associated with questions of the form $S$ and $i$ in $G_{\code,M}$. Using the definition of $\nu$, these satisfy that 
\begin{equation}\label{eq:main-0}
\Es{i\in\{1,\ldots,n\}}\sum_b \big\|P^{W,i}_b - (w^W)^* \tilde{P}_i^W w^W \big\|^2_2 \,=\, O(\delta_1)\;.
\end{equation}
\hnote{What is $\tilde{P}^W_i$, where did that come from?}
Furthermore, since $(\code,M)$ is Abelian, there is a  POVM $\{\tilde{P}^W_u\}_{u\in \field^n}$ such that $\sum_{u\in \code}\tilde{P}^W_u = \Id$ and for each $i\in \{1,\ldots,n\}$, $\tilde{P}^{W,i}_b = \sum_{a:a_i=b} \tilde{P}^W_a$.% where  $u_i = G_\code e_i$.
